Automatically generating alternative hypotheses and controls

### 2018-02-20 topics.tex
initial commit

\usepackage{amsmath}
\usepackage[margin=0.1in]{geometry}
\usepackage{booktabs}
\usepackage{tipa}
\usepackage{amssymb}
\usepackage{tikz}
\usetikzlibrary{fit}
%\usepackage{color}
\usepackage[dvipsnames]{xcolor}
\usepackage{multirow}
\usepackage{hyperref}

@@

\title{Discovering the Rules of language using Bayesian Program Learning}
%\author{Kevin Ellis, Timothy O'Donnell}

\newcommand{\nextForm}[1]{\rotatebox[origin=c]{270}{$_{\curvearrowright}$}$_{#1}$}
\DeclareMathOperator*{\argmin}{arg\,min}

@@

%% How do linguists come up with phonological rules, how do kids learn
%% artificial grammars, and how does one acquire pig latin?  The
%% solutions to these problems share a common representation, which we
%% show can be modeled as a program, and the corresponding learning
%% problems modeled as program induction.  This framing lets us apply
%% ideas from Bayesian Program Learning to induce grammars, which
%% combines program synthesis techniques with a compression-based inductive
%% bias. This lets the models capture phonological phenomena like vowel
%% harmony or stress patterns and learn synthetic grammars used in prior
%% studies of artificial grammar learning.  Going beyond individual
%% grammar learning problems, we consider the problem of jointly
%% inferring many related rule systems. By solving many textbook
%% phonology problems, we can ask the model what kind of inductive bias
%% best explains the attested phenomena.

\begin{abstract}
    %% How do linguists come up with phonological rules? How do kids
    %% learn pig latin?  We develop computational models of these and
    %% other phenomena in language.  The unifying element of the models
    %% is to treat grammars as programs, which lets us apply ideas from
    %% the field of program synthesis to learn grammars. This lets the
    %% models capture phonological phenomena like vowel harmony or stress
    %% patterns and learn synthetic grammars used in prior studies of
    %% rule learning.  Going beyond individual grammar learning problems,
    %% we consider the problem of jointly inferring many related rule
    %% systems. By solving many textbook phonology problems, we can
    %% ask the model what kind of inductive bias best explains the
  %% attested phenomena.
Both children and linguists confront a similar problem of inference:
given utterances produced by speakers, together with aspects of the
meaning of those utterances, discover the grammatical principles that
relate form to meaning. We study this abstract computational
problem within the domain of morphophonology, contributing a computational model
that learns phenomena from many natural languages and generalizes in
humanlike ways from data used in behavioral studies of artificial grammar
learning.

Our work draws on two analogies. The \emph{child-as-linguist analogy}
holds that both children and linguists must solve the same abstract
inductive reasoning problem, even though the nature of the input data and underlying mental algorithms
are surely different in precise detail. Accordingly we isolate the
problem of learning morphophonological systems, and show that a single
solution to this problem can capture both linguistic analyses from
natural languages and infant rule learning of artificial languages.
We adopt the framework of ``Bayesian Program learning'' (BPL) -- in
which learning is formulated a synthesizing a program which compactly
describes the input data. This \emph{learning-as-programming analogy}
lets us exploit recent techniques from the field of program synthesis to
induce morphophonological rules from data.
While \emph{child-as-linguist} poses the
computational problem, \emph{learning-as-programming} offers a
solution.
  \end{abstract}

\section{Introduction}

Both children and linguists confront a similar problem of inference:
given utterances produced by speakers, together with aspects of the
meaning of those utterances, discover the grammatical principles that
relate form to meaning. Here, we study this abstract computational
problem within the domain of morphophonology, the part of language
that builds up words from simpler parts such as stems, prefixes, and
suffixes (morphology) and then specifies how the words are realized as
sound patterns (phonology).  Our contribution is a computational model
that learns phenomena from 26 natural languages and generalizes in
humanlike ways from data used in infant studies of artificial grammar
learning.

Our work draws on two analogies. The \emph{child-as-linguist analogy}
holds that both children and linguists must solve the same abstract
inductive reasoning problem, even though the nature of the input data,
the process of theory construction, and underlying mental algorithms
are surely different in precise detail. Accordingly we isolate the
problem of learning morphophonological systems, and show that a single
solution to this problem can capture both linguistic analyses from
natural languages and infant rule learning of artificial languages.

We adopt the framework of ``Bayesian Program learning'' (BPL) -- in
which learning is formulated a synthesizing a program which compactly
describes the input data. This \emph{learning-as-programming analogy}
explains the flexibility and richness of a grammar
as a consequence of 
While \emph{child-as-linguist} poses the
computational problem, \emph{learning-as-programming} offers a
solution.

We develop a programming language tailored to expressing
morphophonological generalizations.
For children this programming language is innate and is
commonly called ``Universal Grammar'' (UG). Linguists instead must
uncover this programming language by looking at many languages and
searching for commonalities. Learning what makes up UG unfolded over
evolutionary time for children and is ongoing within linguistics.  BPL
reinterprets the problem of discovering UG.  In BPL, UG is the
programming language combined with an inductive bias over programs.
Here we estimate this inductive bias empirically from naturally
occurring morphophonological systems as a way of learning to learn
grammars.

\section{Morphophonological phenomena}

What is phonology supposed to explain? Familiar phenomena from
English help motivate phonological grammars.  The plural of \emph{cat}
is pronounced \textipa{k\ae ts} (``cats''), while the plural of
\emph{dog} is pronounced \textipa{dagz} (``dogs''). A phonological
rule accounts for the alternation between \textipa{s} and \textipa{z}:
\textipa{z} becomes \textipa{s} at the end of words whenever the
preceding sound is unvoiced (does not involve vibrating the vocal
cords). This generalization holds throughout the entirety of English.
The count system of Tibetan presents another puzzle.
``1'' in Tibetan is pronounced
\textipa{\|x{j}ig}, ``10'' is \textipa{\|x{j}u},
yet ``11'' ($ = 10 + 1$) is \textipa{\|x{j}ug\|x{j}ig},
rather than \textipa{\|x{j}u\|x{j}ig}.
The explanation lies in Tibetan phonology,
which deletes word initial consonants followed by another consonant (Table~\ref{Tibetan}).
In this analysis of Tibetan,
``1'' has a latent ``underlying'' representation as \textipa{g\|x{j}ig}
but is pronounced as \textipa{\|x{j}ig}.

The aim of phonology is to explain pronunciations of words in terms of
latent underlying representations (eg, \textipa{g\|x{j}ig} for Tibetan
``1'') and the rules that give observed surface pronunciations from
underlying forms (eg, deleting word initial consonants).
Context-sensitive rewrite systems are a Turing complete representation
that also naturally express phonological
rules~\cite{chomsky1968sound}.  We use context-sensitive rewrites as a
representation for programs that implement a language's phonology.
For example, the rewrite in Table~\ref{Tibetan},
\verb|C|$\to\varnothing /$\verb|#_C|, deletes ($\to\varnothing $)
consonants (\verb|C|) at word boundaries (\verb|#|) when followed by
another consonant (\verb|C|).
Thus phonological rule learners, whether they be children, linguists, or machines, must solve a program induction problem.

\subsection{Inducing Programs}
We model the program learning problem as Bayesian inference, treating the pronunciations of words as observed data and the phonology as a latent variable whose values are programs. So the model's beliefs about possible programs are:
\begin{equation}
  P(\text{program}|\text{words})\propto P(\text{program}|\text{UG})\prod_{w\in\text{words}}\sum_{u\in\text{underlying forms}} P(u)P(w|u,\text{program})
  \end{equation}
Here, Universal Grammar (UG) is an inductive bias over programs. This inductive bias encodes knowledge of the program space (what rules are possible in human language and what rules are impossible), imparting a bias towards more naturally structured rules, like rules written in terms of syllables and consonants.

\subsection{The role and importance of universal grammar}
                We claim any plausible theory of linguistic rule learning needs a universal grammar. Studies of infant grammar learning suggests that babies come equipped with prior knowledge of what language should look like. This prior knowledge constrains the space of allowed generalizations, accounting for the astonishing pace at which children acquire language. This line of reasoning is a ``poverty-of-the-stimulus'' style argument: if children can induce so much grammar from such meager data, something else must bridge the gap. Because universal grammar also constrains the space of allowed grammars, we would expect that humans do not learn certain generalizations, even when the generalizations seem to hold in the data. Indeed this is the case: this ``surfeit-of-the-stimulus'' style argument  explains phenomena in natural languages (for example, statistical regularities that seem to hold in Turkish are not actually used by speakers) and in infant artificial language learning (for example, infants acquire certain phonological generalizations from just a few examples, while never acquiring other generalizations that are seemingly just a simple). While previous accounts of universal grammar were abstract, our model is concrete and computational. We define universal grammar as a probability mass function that puts zero mass on impossible or unattested generalizations, places more mass on programs licensing the generalizations that people learn quickly and naturally, and then smoothly falls off in probability mass as the programs become less natural.

\subsection{Learning natural language phonology}                
Our model learns a diverse set of phonological generalizations found in natural languages. We compiled a multilingual data set of words from standard linguistic textbooks. In these data sets, our model recovers allophonic relationships (e.g., l vs. r in Japanese) and systems that form different inflections of words (e.g., English verb inflections, or the Tibetan count system). Systems like these have received intense study within computational linguistics. But no prior model could explain the diversity of phonological phenomena found even in a textbook. To explain many diverse phenomena we need a rich, flexible representation -- programs -- coupled to a linguisticly informed inductive bias -- universal grammar. Our model learns simple generalizations like English verbal inflections, as well as phenomena like vowel harmony (in the language Kikuria, the word tiga becomes tegera in a different inflection, or ruga becomes rogera), or tonal rules in languages with tone (in the language Kerewe, different inflections cause tones to appear, disappear, or move to different parts of a word).

For these results, the learning-as-programming analogy does the work for us. BPL moves beyond the impoverished representations of prior learners (neural networks as in, or rigid classes of symbolic rules as in) instead represented generalizations as programs. We claim that programs are a sufficiently general representation to support the kinds of rules found in phonology.

\begin{table}\centering
\begin{tabular*}{10cm}{lcl}
  \toprule
  Example data&Phonology\\\midrule
\begin{tabular}{ll}
1&\textipa{\|x{j}ig}\\
4&\textipa{\|x{s}i}\\
5&\textipa{Na}\\
9&\textipa{gu}\\
10&\textipa{\|x{j}u}\\
11 ($= 10 + 1$) & \textipa{\|x{j}ug\|x{j}ig}\\
14 ($= 10 + 4$) & \textipa{\|x{j}ub\|x{s}i}\\
15 ($= 10 + 5$) & \textipa{\|x{j}uNa}\\
19 ($= 10 + 9$) & \textipa{\|x{j}urgu}\\
40 ($= 4 + 10$) & \textipa{\|x{s}ib\|x{j}u}\\
50 ($= 5 + 10$) & \textipa{Nab\|x{j}u}\\
90 ($= 9 + 10$) & \textipa{gub\|x{j}u}
\end{tabular}
&
\begin{tabular}{l}
  \emph{Consonant cluster reduction:}\\
  \verb|C|$\to\varnothing $ $/$ \verb|# _ C|\\
  \end{tabular}

  \\
  \bottomrule  \end{tabular*}
\caption{Tibetan count system.}
\label{Tibetan}
  \end{table}

\pagebreak

%% 2 insights: analogy between children and linguists. analogy between
%% grammars and programs.

%% The core problem we're pushing on is \textbf{structure learning in
%%   language}. It makes sense that you could learn parameters and
%% lexical items, but could you learn something with more the structure
%% of a grammar? Here we answer that question in the affirmative within a
%% restricted domain (phonology) on restricted problems (smaller data
%% sets and simpler sets of rules).

%% Why focus on some problems that feel more like computational linguistics and other problems that feel more like developmental psychology? Connection to child-as-linguist~\cite{chomsky1988current}, child-as-scientist, child-as-hacker\cite{Rule2015RepresentingAL}.

%% What is the connection between what children do and what linguists do? From the point of view of this research project,
%% the difference lies in the nature and role of universal grammar.
%% For a linguist, universal grammar is an unknown to be discovered.
%% When analyzing a new data set, a linguist might need to entertain a broader or more expressive universal grammar.
%% For the child, universal grammar is an innate part of our genetic endowment.

\section{The Model}

\begin{figure}[h!]\centering
  \includegraphics[width = 0.9\textwidth]{generativeModel.pdf}
  \caption{{\color{red}Red}: Both the morphology $\mathcal{M}$ and
    phonology $\mathcal{P}$ are programs either build through
    concatenation or function composition
    (Tables~\ref{morphologyGrammar} \&~\ref{phonologyGrammar}). These programs explain how words are built in both natural and artificial grammars ({\color{blue}blue}). Learners observe pronunciations of words ({\color{magenta}}) in order to recover these programs, but with the aid of a strong inductive bias (Universal Grammar, {\color{purple}purple})}.
  \label{beautiful}
\end{figure}

\section{Artificial Grammar Learning}

The fundamental hypothesis underlying AGL research is that
artificial grammar learning must engage some shared resource with first language acquisition --
and so we can gain insights into first language acquisition by
instead studying AGL in controlled experiments.

Our contribution to AGL research is the demonstration that
a single BPL model can explain (1) that infants quickly acquire
certain simple grammars very easily, and at the same time (2)
those same infants go on to learn
the very complicated grammar of their first language.
We collected grammars from AGL datasets~\cite{gerken2010infants,marcus1999rule,frank2011three}
and presented our BPL learner with data drawn from these grammars (Table~\ref{artificialGrammarTable}),
finding that our model recovers the generative processes that
 underlies these data sets.
\begin{table}[h!]
\centering\begin{tabular}{lcll}
  \toprule
  Grammar&Example input to learner&MAP grammar&Natural language analogues\\\midrule
  \begin{tabular}{l}
    ABA (same/different/same)\\
  Marcus et al. 1999.     
    \end{tabular}&
  \begin{tabular}{l}
    wofewo\\
    lovilo\\
    fimufi
  \end{tabular}&
  \begin{tabular}{l}
$\varnothing\to \sigma_i$\verb| / #_|$\sigma \sigma_i$
  \end{tabular}& Reduplication (eg, Tagalog)
  \\\midrule
  \begin{tabular}{l}
      ABB (different/same/same)\\
  Marcus et al. 1999. 
    \end{tabular}&
  \begin{tabular}{l}
        wofefe\\
    lovivi\\
    fimumu
  \end{tabular}&
  \begin{tabular}{l}
$\varnothing\to \sigma_i$\verb| / |$\sigma_i$\verb|_#|
  \end{tabular}&
  Reduplication (eg, Tagalog)\\\midrule
  \begin{tabular}{l}
      ABx (different/different/constant)
    \end{tabular}&
  \begin{tabular}{l}
        wofeka\\
    lovika\\
    fimuka
  \end{tabular}&
  \begin{tabular}{l}
stem$ + x$
  \end{tabular}&concatenative morphology
  \\\midrule
  \begin{tabular}{l}
    AAx (same/same/constant)\\Gerken 2006. %~\cite{gerken2010infants}
    \end{tabular}&
  \begin{tabular}{l}
        wowoka\\
    loloka\\
    fifika
  \end{tabular}&
  \begin{tabular}{l}
    stem + $x$\\
$\varnothing\to \sigma_i$\verb| / #_|$\sigma_i$
  \end{tabular}&
  \begin{tabular}{l}
    reduplication\\concatenative morphology
    \end{tabular}\\\midrule
  \begin{tabular}{l}
    AxA (same/constant/same)\\Gerken 2006. %~\cite{gerken2010infants}
    \end{tabular}&
  \begin{tabular}{l}
        wokawo\\
    lokalo\\
    fikafi
  \end{tabular}&
  \begin{tabular}{l}
$x + $ stem\\
    $\varnothing$ $\to\sigma_i$  / \verb|# _| $\sigma$ $\sigma_i$
  \end{tabular}&
  \begin{tabular}{l}
    Infixing\\Reduplication
    \end{tabular}
  \\\midrule
  \begin{tabular}{l}
    Pig Latin
    \end{tabular}&
  \begin{tabular}{l}
    \textipa{pIg}$\to$\textipa{Igpe}\\
    \textipa{latIn}$\to$\textipa{atIle}\\
    \textipa{\ae sk}$\to$\textipa{\ae ske}\\
  \end{tabular}&
  \begin{tabular}{l}
      $\varnothing\to$\verb|C|$_i$ $/$ \verb|#C|$_i$\verb| [ ]* _ #|\\%
      $\varnothing\to$ \textipa{e} $/$ \verb| _ #|\\%
 \verb|C|$\to\varnothing$ $/$ \verb|# _|
  \end{tabular}  & \begin{tabular}{l}
    \textbf{Tim, is there an  analog?}\\
    \textbf{word initial segments moves to  }\\
    \textbf{word final?}
    \end{tabular}
  \bottomrule  \end{tabular}
\caption{Using BPL to model artificial grammar learning. Learner given 5 examples.}\label{artificialGrammarTable}
\end{table}

Just like in first language acquisition,
these AGL learning setups introduce a trade-off between
grammars that are likely and therefore small (``parsimony'')  and grammars that
best fit the data.
For example, a learner could infer a grammar that just memorized to the data (perfect fit but poor parsimony)
or it could infer a grammar that can generate every
possible word (parsimonious but a poor fit).
Effective grammar learners must navigate this trade-off.

@@

  \caption{Pareto fronts for ABB  (Marcus~1999; top two) \& AAX  (Gerken~2006; bottom two) learning problems fo either one example word (left column) or three example words  (right column). {\color{red}Red shade}: ground truth grammar. {\color{red!60}Pink shade}: shares structure with ground truth grammar. White shade: incorrect generalizations. As the number of examples increases, the Pareto fronts develop a sharp kink around the ground truth grammar, which indicates a stronger preference for the correct grammar. With one example the kinks can still exists but are less pronounced.}\label{quadrupleFrontier}
\end{figure}

@@

A successful model of
artificial grammar learning should also explain at least some aspects of first language acquisition.
As a first step in this direction,
we take the same model that we applied to
these AGL learning problems and use it to solve phonology textbook exercises.
Each of these textbook exercises is an inductive reasoning
problem where the learner must
explain surface pronunciations
in terms of a latent generative model (a program).
Figure~\ref{naturalLanguageExample}
shows our model solving a
representative textbook problem.

\textbf{TIM: Should I talk about UG? We needed it for Tibetan. We probably don't have space.}

\begin{figure}\centering
  \begin{minipage}{7cm}\centering
    \textbf{Textbook Tibetan counting problem:}
    \begin{tabular}{ll}\toprule
Number & Surface form
\\ \midrule
10 & \textipa{d\super Zu}\\
1 & \textipa{d\super Zig}\\
11 & \textipa{d\super Zugd\super Zig}\\
4 & \textipa{Si}\\
%14 & \textipa{d\super ZubSi}\\
40 & \textipa{Sibd\super Zu}\\
9 & \textipa{gu}\\
19 & \textipa{d\super Zurgu}\\
%90 & \textipa{gubd\super Zu}\\
5 & \textipa{Na}\\
%15 & \textipa{d\super ZuNa}\\
%50 & \textipa{Nabd\super Zu}\\
\bottomrule
\end{tabular}
  \end{minipage}%
  \hspace{0.5cm}%
  \begin{minipage}{6cm}\centering
    \textbf{Our model's solution:}

@@

    Induced underlying representations:
\\\begin{tabular}{ll}\toprule
1 & \textipa{gd\super Zig}\\
4 & \textipa{bSi}\\
5 & \textipa{Na}\\
9 & \textipa{rgu}\\
10 & \textipa{bd\super Zu}
\bottomrule
\end{tabular}    
  \end{minipage}
  \caption{Example textbook phonology exercise (left) and the solution our model discovers for it (right). To solve this problem the model must jointly infer unobserved underlying representations for each of the Tibetan numbers, along with a phonological rule that explains their surface pronunciations. The same code that calculated the Pareto fronts (Figure~\ref{quadrupleFrontier}) produces this analysis of the Tibetan count system.}\label{naturalLanguageExample}
  \end{figure}

@@

\pagebreak
\section{Appendix}

\begin{table*}[h]
\centering
\begin{tabular*}{\textwidth}{ll}
\toprule
Grammar rule  & English description \\\midrule
$\mathcal{P}\to \mathcal{R}\circ \mathcal{R}\circ \cdots \circ \mathcal{R}$&Phonology $\mathcal{P}$ is a sequence of compositions of rewrites $\mathcal{R}$\\
$\mathcal{R}\to \mathcal{F}\longrightarrow \mathcal{C} / \mathcal{T}\_\mathcal{T}$& Rewrite focus $\mathcal{F}$ to structural change $\mathcal{C}$ between triggers $\mathcal{T}$\\
$\mathcal{T}\to \# \mathcal{T}' | \mathcal{T}'$ & Triggers optionally match end of string\\
$\mathcal{T}'\to \epsilon | \mathcal{X} \mathcal{T}' | \mathcal{X}^* \mathcal{T}'$ & Triggers are sequences of matrices $\mathcal{X}$ possibly with Kleene star\\
$\mathcal{X}\to a|t|s|\cdots$&Matrices can be constant phonemes\\
$\mathcal{X}\to [\pm \mathcal{G} \; \pm \mathcal{G} \; \cdots \; \pm \mathcal{G}]$& Matrices check features $\mathcal{G}$\\
$\mathcal{X}\to \sigma$&Matrices can be whole syllables\\
$\mathcal{G}\to \text{voice}|\text{nasal}|\cdots$&Standard phonological features\\
$\mathcal{F}\to \mathcal{X}$&Focus can be a feature matrix\\
$\mathcal{F}\to \varnothing$&Focus can be empty (insertion rule)\\
$\mathcal{F}\to\mathbb{Z}$&Focus can be one of the triggers (copies it)\\
$\mathcal{C}\to\mathcal{X}$&Structural change can be a feature matrix\\
$\mathcal{C}\to\varnothing$&Deletion rule\\
\bottomrule
\end{tabular*}
\caption{Grammar over phonologies. The non-terminal $\mathcal{P}$ is the start symbol for the grammar.}
\label{phonologyGrammar}
\end{table*}

\begin{table*}[h]
\centering
\begin{tabular*}{\textwidth}{ll}
\toprule
Grammar rule  & English description \\\midrule
$\mathcal{W}\to (\mathcal{M}) \mathcal{M} (\mathcal{M})$&Morphology builds words $\mathcal{W}$ by concatenating morphemes $\mathcal{M}$. Prefix/Suffix optional (in parentheses)\\
$\mathcal{M}\to$ sequence of phonemes\bottomrule
\end{tabular*}
\caption{Grammar over morphologies. The non-terminal $\mathcal{W}$ is the start symbol for the grammar.}
\label{morphologyGrammar}
\end{table*}

\pagebreak
\subsection{Inference}

The solver takes as input:
\begin{itemize}
\item A set of inflected surface forms, $\{x_n\}$. Here the variable $x_n$ refers to all of the inflections of a word. So an example set of 3 inflected surface forms might be $\{(\text{fish},\text{fished}), (\text{walk},\text{walked}), (\text{eat},\text{ate})\}$.
\item A bound on the number of rules to consider synthesizing, $B_R$.
\end{itemize}
It produces as output:
\begin{itemize}
\item A set of underlying forms, $\{u_n\}$.
\item A morphology consisting of a prefix ($\{p_i\}$) and a suffix ($\{s_i\}$)  for each inflection
\item  A sequence of rules, $\{r_j\}_{j = 1}^{B_R}$
\end{itemize}
It is \emph{guaranteed} to discover the rules, morphology, and underlying forms minimizing:
\begin{equation}
  \text{solver}(\{x_n\},B_R) = \argmin_{\{r_j\}_{j = 1}^{B_R}, \{u_n\}, \{p_i\}, \{s_i\}} \sum_j \text{cost}(r_j) + \sum_i (\text{cost}(p_i) + \text{cost}(s_i) ) + \sum_n \text{cost}(u_n)
\end{equation}
In exchange for this strong guarantee, however, it has only a very
weak guarantee on the amount of time it will take to discover the
solution: in the worst case, it will take time exponential size of the
problem. In practice it takes close to a day for it to do exact search over 3 rules.
For easy problems, like the AGL experiments, I can get away with 4 rules if I wait for a few hours.

One can also ask the solver for the top $K$ solutions minimizing the cost function, which I will write:
\begin{equation}
  \text{solver}_K(\{x_n\},B_R) = \text{the }K\text{ minimum cost solutions.}
\end{equation}
One can also provide additional constraints to the solver, for example telling it what the underlying forms are,
or telling it what some of the rules are. I will write these constraints by adding an extra argument to the solver
that looks like conditioning in Bayesian inference. For example,
\begin{equation}
  \text{solver}_K(\{x_n\},B_R|u_1 = \text{fish}) = \text{the }K\text{ minimum cost solutions where }u_1\text{ has to be fish.}
  \end{equation}

>>>>

\usepackage{amsmath}
\usepackage[margin=0.1in]{geometry}
\usepackage{booktabs}
\usepackage{tipa}
\usepackage{amssymb}
\usepackage{dsfont}
\usepackage{tikz}
\usetikzlibrary{fit}
%\usepackage{color}
\usepackage[dvipsnames]{xcolor}
\usepackage{multirow}
\usepackage{hyperref}
\usepackage{amsfonts}       % blackboard math symbols

@@

\title{Automatically generating alternative hypotheses and controls}
%\author{Kevin Ellis, Timothy O'Donnell}

\newcommand{\nextForm}[1]{\rotatebox[origin=c]{270}{$_{\curvearrowright}$}$_{#1}$}
\DeclareMathOperator*{\argmin}{arg\,min}
\newcommand{\Expect}{\mathds{E}} %{{\rm I\kern-.3em E}}
\newcommand{\indicator}{\mathds{1}} %{{\rm I\kern-.3em E}}
\newcommand{\expect}{\mathds{E}} %{{\rm I\kern-.3em E}}
\newcommand{\probability}{\mathds{P}} %{{\rm I\kern-.3em P}}

@@

\section{Automatically generating grammars with Bayesian program learning}

\subsection{A Probabilistic Framing}

The goal of the grammar learner, whether it is a child, linguist, or computer program, is to
infer the grammar that gave rise to
a collection of utterances.
One way of
modeling this inference problem is to (1) place a prior distribution over grammars, for example, a distribution that puts more probability mass on shorter or simpler grammars; (2) equip each grammar with a
\emph{likelihood model} that specifies exactly how likely a grammar is to produce a given utterance;
and then (3) use Bayes's rule to work backwards from the
utterances to the grammar that was likely to produce them:
\begin{equation}
  \probability\left[\text{grammar}|\text{utterances} \right]\propto\underbrace{\probability\left[\text{utterances}|\text{grammar} \right]}_{\text{Likelihood}}\times\underbrace{\probability\left[\text{grammar} \right]}_{\text{Prior}}\label{posterior}
\end{equation}
We can think of the above equation as a recipe for
scoring how well a grammar explains a
collection of utterances.
Once we have a prior and a likelihood,
we know how to evaluate how well a grammar explains the data, modulo the assumptions implicit in the
prior and likelihood.

A simple and intuitive prior over grammars is one which penalizes longer grammars and favors parsimonious grammars, for example defining $\probability\left[\text{grammar} \right]\propto\exp\left(- \text{\# symbols in grammar} \right)$. An example of a likelihood model,
and the one we use here,
is just to count the number of extra bits or symbols needed to encode an utterance given the grammar\footnote{This is equivalent to a minimum description length (MDL) interpretation}.

Now in theory this probabilistic framing
has now bought us a procedure for automatically discovering
grammars that do a good job explaining the data.
In practice,
the space of all grammars is infinite and combinatorial,
and so cashing out this framing requires a
efficient procedure for searching the space of grammars and honing in on those that maximize Equation~\ref{posterior}.
A subfield of computer science, called \textbf{program synthesis},
has developed efficient techniques for
solving seemingly intractable combinatorial search problems
like those that arise in grammar learning.
We use a program synthesis tool called Sketch~\cite{solar2008program}
to search for grammars maximizing~\ref{posterior},
which was previously applied to language learning problems in~\cite{ellis2015unsupervised}.
For more detail on these program synthesis techniques we refer the reader to~\cite{solar2008program}.
This cashing out of the framing makes
our approach an instance of  \textbf{Bayesian Program Learning (BPL)}; see~\cite{lake2015human}.

\subsection{Modeling solutions to AGL problems}

The fundamental hypothesis underlying AGL research is that
artificial grammar learning must engage some shared resource with first language acquisition --
and so we can gain insights into first language acquisition by
instead studying AGL in controlled experiments.
Our contribution here is the demonstration that
a single BPL model can explain how a learner can quickly infer a grammar represented as SPE--style rules.
We collected grammars from AGL datasets~\cite{gerken2010infants,marcus1999rule,frank2011three}
and presented our BPL learner with data drawn from these grammars (Table~\ref{artificialGrammarTable}),
finding that our model automatically discovers the grammars that
 underly these data sets.
\begin{table}[h!]
\centering\begin{tabular}{lcll}
  \toprule
  Grammar&Example input to learner&MAP grammar&Natural language analogues\\\midrule
  \begin{tabular}{l}
    ABA (same/different/same)\\
  Marcus et al. 1999.     
    \end{tabular}&
  \begin{tabular}{l}
    wofewo\\
    lovilo\\
    fimufi
  \end{tabular}&
  \begin{tabular}{l}
$\varnothing\to \sigma_i$\verb| / #_|$\sigma \sigma_i$
  \end{tabular}& Reduplication (eg, Tagalog)
  \\\midrule
  \begin{tabular}{l}
      ABB (different/same/same)\\
  Marcus et al. 1999. 
    \end{tabular}&
  \begin{tabular}{l}
        wofefe\\
    lovivi\\
    fimumu
  \end{tabular}&
  \begin{tabular}{l}
$\varnothing\to \sigma_i$\verb| / |$\sigma_i$\verb|_#|
  \end{tabular}&
  Reduplication (eg, Tagalog)\\\midrule
  \begin{tabular}{l}
      ABx (different/different/constant)
    \end{tabular}&
  \begin{tabular}{l}
        wofeka\\
    lovika\\
    fimuka
  \end{tabular}&
  \begin{tabular}{l}
stem$ + x$
  \end{tabular}&concatenative morphology
  \\\midrule
  \begin{tabular}{l}
    AAx (same/same/constant)\\Gerken 2006. %~\cite{gerken2010infants}
    \end{tabular}&
  \begin{tabular}{l}
        wowoka\\
    loloka\\
    fifika
  \end{tabular}&
  \begin{tabular}{l}
    stem + $x$\\
$\varnothing\to \sigma_i$\verb| / #_|$\sigma_i$
  \end{tabular}&
  \begin{tabular}{l}
    reduplication\\concatenative morphology
    \end{tabular}\\\midrule
  \begin{tabular}{l}
    AxA (same/constant/same)\\Gerken 2006. %~\cite{gerken2010infants}
    \end{tabular}&
  \begin{tabular}{l}
        wokawo\\
    lokalo\\
    fikafi
  \end{tabular}&
  \begin{tabular}{l}
$x + $ stem\\
    $\varnothing$ $\to\sigma_i$  / \verb|# _| $\sigma$ $\sigma_i$
  \end{tabular}&
  \begin{tabular}{l}
    Infixing\\Reduplication
    \end{tabular}
  \\\midrule
  \begin{tabular}{l}
    Pig Latin
    \end{tabular}&
  \begin{tabular}{l}
    \textipa{pIg}$\to$\textipa{Igpe}\\
    \textipa{latIn}$\to$\textipa{atIle}\\
    \textipa{\ae sk}$\to$\textipa{\ae ske}\\
  \end{tabular}&
  \begin{tabular}{l}
      $\varnothing\to$\verb|C|$_i$ $/$ \verb|#C|$_i$\verb| [ ]* _ #|\\%
      $\varnothing\to$ \textipa{e} $/$ \verb| _ #|\\%
 \verb|C|$\to\varnothing$ $/$ \verb|# _|
  \end{tabular}  & \begin{tabular}{l}
    \textbf{Tim, is there an  analog?}\\
    \textbf{word initial segments moves to  }\\
    \textbf{word final?}
    \end{tabular}
  \bottomrule  \end{tabular}
\caption{Using BPL to model artificial grammar learning. Learner given 5 examples.}\label{artificialGrammarTable}
\end{table}

\subsection{Automatically exploring the space of alternative hypotheses}

Just like in first language acquisition,
these AGL learning setups introduce a trade-off between
grammars that are likely under the prior (and are therefore small; ``parsimony'')  and grammars that
assign high likelihood to the data (and therefore closely ``fit the data'', having high likelihood).
For example, a learner could infer a grammar that just memorized to the data (perfect fit but poor parsimony)
or it could infer a grammar that can generate every
possible word (parsimonious but a poor fit).
Effective grammar learners must navigate this trade-off.

@@

  \caption{Pareto fronts for ABB  (Marcus~1999; top two) \& AAX  (Gerken~2006; bottom two) learning problems fo either one example word (left column) or three example words  (right column). Rightward on x-axis corresponds to more parsimonious grammars and upward on y-axis corresponds to grammars that best fit the data, so the best grammars live in the upper right corners of the graphs. {\color{red}Red shade}: ground truth grammar. {\color{red!60}Pink shade}: shares structure with ground truth grammar. White shade: incorrect generalizations. As the number of examples increases, the Pareto fronts develop a sharp kink around the ground truth grammar, which indicates a stronger preference for the correct grammar. With one example the kinks can still exists but are less pronounced.}\label{quadrupleFrontier}
\end{figure}

@@

A successful model of artificial grammar learning should also explain
at least some aspects of first language acquisition.  As a first step
in this direction, we take the same model that we applied to these AGL
learning problems and use it to solve phonology textbook exercises.
Each of these textbook exercises is an inductive reasoning problem
where the learner must explain surface pronunciations in terms of a
latent generative model (a grammar).
Figure~\ref{naturalLanguageExample} shows our model solving a
representative textbook problem.  We believe this extension two
natural language is an important next step for models of artificial
grammar learning,
and helps deliver on the basic promises and presuppositions of AGL research.

\begin{figure}[H]\centering
  \begin{minipage}{7cm}\centering
    \textbf{Textbook Tibetan counting problem:}
    \begin{tabular}{ll}\toprule
Number & Surface form
\\ \midrule
10 & \textipa{d\super Zu}\\
1 & \textipa{d\super Zig}\\
11 & \textipa{d\super Zugd\super Zig}\\
4 & \textipa{Si}\\
%14 & \textipa{d\super ZubSi}\\
40 & \textipa{Sibd\super Zu}\\
9 & \textipa{gu}\\
19 & \textipa{d\super Zurgu}\\
%90 & \textipa{gubd\super Zu}\\
5 & \textipa{Na}\\
%15 & \textipa{d\super ZuNa}\\
%50 & \textipa{Nabd\super Zu}\\
\bottomrule
\end{tabular}
  \end{minipage}%
  \hspace{0.5cm}%
  \begin{minipage}{6cm}\centering
    \textbf{Our model's solution:}

@@

    Induced underlying representations:
\\\begin{tabular}{ll}\toprule
1 & \textipa{gd\super Zig}\\
4 & \textipa{bSi}\\
5 & \textipa{Na}\\
9 & \textipa{rgu}\\
10 & \textipa{bd\super Zu}
\bottomrule
\end{tabular}    
  \end{minipage}
  \caption{Example textbook phonology exercise (left, taken from~\cite{9780511808869}) and the solution our model discovers for it (right). To solve this problem the model must jointly infer unobserved underlying representations for each of the Tibetan numbers, along with a phonological rule that explains their surface pronunciations. The same code that calculated the Pareto fronts (Figure~\ref{quadrupleFrontier}) produces this analysis of the Tibetan count system.}\label{naturalLanguageExample}
  \end{figure}

@@
